---
title: "Static Temperature Data"
author: "Brittnany Graham"
editor: visual
format:
  html:
    embed-resources: false
execute:
  echo: false
  message: false
  warning: false
  results: 'hide'
---

## River Temperature Plots

::: callout-note
## This website still updates as data comes in. This is to reflect and document the duration of my study, so this will continuously change as I gather more data! :)
:::

I belive it's important to connect the community with science everyday, even in small ways!
This page is dedicated to some temperature data I've collected thus far in my studies!
It's so interesting to look at and I try to make them so folks can understand what's happening!
Plus, I find it enjoyable being able to show people how cool science can be!

```{r}
library(tidyverse)
library(ggpubr)
library(stringr)
library(viridis)
```

```{r}
temp_csv_files <- list.files(path = "C:/Users/bgraha33/OneDrive - University of Tennessee/Desktop/New Dataset/Temperature CSV Files", pattern = '\\.csv$', full.names = TRUE)


temp_list <- map(temp_csv_files, read.csv)
combined_temp_data <- bind_rows(temp_list)


temp <- combined_temp_data

#converting it to Fahrenheit

temp <- temp %>%
  mutate(TemperatureF = TemperatureC * 9/5 + 32)

# Rounding up to the nearest whole number

temp <- temp %>%
  mutate(TemperatureF = round(TemperatureF))


## Clinch River

## Receiver 802012

clinch_river_temp <- temp %>%
  filter(ReceiverNumber == "802012")

clinch_river_temp$DateAndTimeUTC <- mdy_hm(clinch_river_temp$DateAndTimeUTC)
clinch_river_temp$Date <- mdy(clinch_river_temp$Date)

```

## Emory River Temperature

::: callout-important
## The Emory River experience record highs and record low river levels during 2025
:::

```{r}
emory_temp <- temp %>%
  filter(ReceiverNumber == "802023")

# Changed to posixct

emory_temp$DateAndTimeUTC <- mdy_hm(emory_temp$DateAndTimeUTC)

emory_temp$Date <- mdy(emory_temp$Date)

## Graphing it now

emory_static_plot <- ggplot(emory_temp, aes(x = Date, y = TemperatureF)) +
  annotate("rect",
           xmin = min(emory_temp$Date),
           xmax = max(emory_temp$Date),
           ymin = 64.5,
           ymax = 72,
           fill = "#FDB736",
           alpha = 0.3) +
  annotate("label",
           x = as.Date("2025-04-25"),
           y = 85,
           label = "Lethal Thermal Limit (27°C)",
           fill = "white",
           size = 4) +
  annotate("label",
           x = as.Date("2025-12-25"),
           y = 70,
           label = "Optimal Growth Range",
           fill = "white",
           size = 4) +
  geom_line(aes(color = TemperatureF), linewidth = 1) +
  geom_hline(yintercept = 81,
             color = "#D55E00",
             linetype = "dashed",
             linewidth = 1) +
  scale_color_viridis(option = "C") +
  scale_x_date(date_breaks = "1 month",
               date_labels = "%b \n%Y") +
  scale_y_continuous(breaks = seq(0, 90, 5)) +
  labs(title = "Emory River Daily Temperature",
       x = "Month",
       y = "Temperature (°F)") +
  theme_classic(base_size = 14) +
  theme(plot.title = element_text(size = 16, face = "bold"),
        legend.position = "none",
        axis.title.x = element_text(face = "bold", size = 12),
        axis.title.y = element_text(face = "bold", size = 12),
        axis.text.x = element_text(face = "bold", size = 10),
        axis.text.y = element_text(face = "bold", size = 10),)

print(emory_static_plot)
```

## Temperature of Tennessee River - Near Browder Shoals

The area is also known as Browder Bend or colloquially understood as the area near the A-Frame house.

The temperature in this area doesn't quite reach the lethal limit, but toes the line during the summer.

```{r}
tn_river <- temp %>%
  filter(ReceiverNumber == "802025")

tn_river$Date <- mdy(tn_river$Date)


## Graphing
tn_plot <- ggplot(tn_river, aes(x = Date, y = TemperatureF)) +
  annotate("rect",
           xmin = min(tn_river$Date),
           xmax = max(tn_river$Date),
           ymin = 64.5,
           ymax = 72,
           fill = "#FDB736",
           alpha = 0.3) +
  annotate("label",
           x = as.Date("2025-04-25"),
           y = 85,
           label = "Lethal Thermal Limit (27°C)",
           fill = "white",
           size = 4) +
  annotate("label",
           x = as.Date("2025-07-20"),
           y = 70,
           label = "Optimal Growth Range",
           fill = "white",
           size = 4) +
  geom_line(aes(color = TemperatureF), linewidth = 1) +
  geom_hline(yintercept = 81,
             color = "#D55E00",
             linetype = "dashed",
             linewidth = 1) +
  scale_color_viridis(option = "C") +
  scale_x_date(date_breaks = "1 month",
               date_labels = "%b \n%Y") +
  scale_y_continuous(limits = c(30, 90),
                     breaks = seq(30, 90, 5))+ 
  labs(title = "Tennessee River Daily Temperature",
       x = "Month",
       y = "Temperature (°F)") +
  theme_classic(base_size = 14) +
  theme(plot.title = element_text(size = 16, face = "bold"),
        legend.position = "none",
        axis.title.x = element_text(face = "bold", size = 12),
        axis.title.y = element_text(face = "bold", size = 12),
        axis.text.x = element_text(face = "bold", size = 10),
        axis.text.y = element_text(face = "bold", size = 10),)

tn_plot
```

## Clinch River Temperature

You can look at the monthly temperature of the Clinch River from April 2025 - February 2026!

The Clinch stays relatively cool during the hottest times during the summer!
Which also makes it ideal for temps for paddling, kayaking, and swimming!

```{r}
clinch_river_temp <- temp %>%
  filter(ReceiverNumber == "802012")

clinch_river_temp$DateAndTimeUTC <- mdy_hm(clinch_river_temp$DateAndTimeUTC)
clinch_river_temp$Date <- mdy(clinch_river_temp$Date)

# str(clinch_river_temp)

## Graphing it
clinch_plot <- ggplot(clinch_river_temp, aes(x = Date, y = TemperatureF)) +
  annotate("rect",
           xmin = min(clinch_river_temp$Date),
           xmax = max(clinch_river_temp$Date),
           ymin = 64.5,
           ymax = 72,
           fill = "#FDB736",
           alpha = 0.3) +
  annotate("label",
           x = as.Date("2025-04-25"),
           y = 85,
           label = "Lethal Thermal Limit (27°C)",
           fill = "white",
           size = 4) +
  annotate("label",
           x = as.Date("2026-01-07"),
           y = 70,
           label = "Optimal Growth Range",
           fill = "white",
           size = 4) +
  geom_line(aes(color = TemperatureF), linewidth = 1) +
  geom_hline(yintercept = 81,
             color = "#D55E00",
             linetype = "dashed",
             linewidth = 1) +
  scale_color_viridis(option = "C") +
  scale_x_date(date_breaks = "1 month",
               date_labels = "%b \n%Y") +
  scale_y_continuous(limits = c(30, 90),
                     breaks = seq(30, 90, 5))+ 
  labs(title = "Clinch River Daily Temperature",
       x = "Month",
       y = "Temperature (°F)") +
  theme_classic(base_size = 14) +
  theme(plot.title = element_text(size = 16, face = "bold"),
        legend.position = "none",
        axis.title.x = element_text(face = "bold", size = 12),
        axis.title.y = element_text(face = "bold", size = 12),
        axis.text.x = element_text(face = "bold", size = 10),
        axis.text.y = element_text(face = "bold", size = 10),)

clinch_plot

```

Again, these are all preliminary temperature data.
I will have at least a complete year of data when I go back to the field and download more data from the receivers to update this page!
